\hydrogen{} and \carbon{} Network Strategies for the European Energy System

### 2024-08-28 report/submissions/journal/paper/paper.tex
add policy conclusions and brief references to recent COM and JRC
CMS work

% Recently, policymakers in Europe have started developing hydrogen and carbon management strategies,
% ~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
In line with the climate neutrality target by 2050 of the European Union (EU) under the European Green Deal~\cite{europeangreendeal}, numerous programs, financial models, and initiatives have been launched to support the ramp-up of the hydrogen and carbon economy. The Renewable Energy Directive III~\cite{DirectiveEU20232023} aims to produce 10~Mt hydrogen from renewable energy sources domestically in the EU. The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt per year by 2030, which will be sequestered within the European Union. The European Innovation Fund~\cite{europeaninnovationfund} supports carbon management technologies to enhance the global competitiveness of European industries.

@@

In the literature, the two network approaches and underlying technologies have been discussed in several publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}. Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions. Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments. The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the option to transport carbon across regions.
Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment. This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway. However, this limited spatial scope fails to consider other sequestration sites and the co-benefits from connecting the \carbon{} network to neighboring countries.
Other high-resolution energy system models do not feature detailed CU technologies and \carbon{} transport. Euro-Calliope, for example, does not feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.

@@

>>>>

% Recently, policymakers in Europe have started developing hydrogen and carbon management strategies,
% ~\cite{GermanyDevelopingStrategy2023,CarbonManagementStrategie}, planning infrastructure components~\cite{CONetz}, and committing to the first carbon utilization projects~\cite{EFuelsPilotPlant2022,OrstedAssumesFull,GROUNDBREAKINGEFUELPRODUCTION,DLREfuelsDLR}.
In line with the climate neutrality target by 2050 of the European Union (EU) under the European Green Deal~\cite{europeangreendeal}, numerous programs, financial models, and initiatives have been launched to support the ramp-up of the hydrogen and carbon economy. The Renewable Energy Directive III~\cite{DirectiveEU20232023} aims to produce 10~Mt hydrogen from renewable energy sources domestically in the EU. The Net-Zero Industry Act~\cite{eu2023netzero} proposes a \carbon{} injection target of 50~Mt per year by 2030, which will be sequestered within the European Union. The European Commission's 2024 industrial carbon management strategy~\cite{comcms} plans for capture of 450~Mt/a by 2050 in the EU, of which almost 250~Mt/a is sequestered.  The European Innovation Fund~\cite{europeaninnovationfund} supports carbon management technologies to enhance the global competitiveness of European industries.

@@

In the literature, the two network approaches and underlying technologies have been discussed in several publications, all of which, however, dealt with the isolated aspects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022,neumannBenefitsHydrogenNetwork2022}. Such techno-economic models, in contrast for example to Integrated Assessment Models, can account for the spatial distribution of carbon sources and sinks, which are crucial to provide a holistic view of the energy system and its technological interactions. Neumann et al.~\cite{neumannBenefitsHydrogenNetwork2022} demonstrated the interchangeability of hydrogen and electricity grid expansion in a climate-neutral European energy system at the expense of higher investments. The underlying, highly-resolved model encompasses the most relevant sectors, however it neglects the option to transport carbon across regions.
Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment. This limited sectoral scope cannot capture important dynamics of carbon management, since it neglects sectors like industry that will need to handle most \carbon{} in the future.
The European Commission's Joint Research Center analysed the possible topology of a \carbon{} network in 2024 \cite{jrc2024}, but didn't include the interaction of  \carbon{} sources and sinks with the rest of the energy system.
A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connected to a remote sequestration site in Norway. However, this limited spatial scope fails to consider other sequestration sites and the co-benefits from connecting the \carbon{} network to neighboring countries.
Other high-resolution energy system models do not feature detailed CU technologies and \carbon{} transport. Euro-Calliope, for example, does not feature carbon nor hydrogen transport, and thus forces generic synthetic fuel production to rely on captive electrolysis and DAC in the same region~\cite{pickeringDiversityOptionsEliminate2022}.

@@

For policy makers these results show the need for coordinated planning across sectors like carbon, hydrogen and synthetic fuels since they are strongly interacting, as well as across borders because of the localized nature of good renewable resources and sequestration potentials. Deploying multiple networks offers some robustness should unforeseen problems arise with one of them, and our results show that the system can even cope with neither a \carbon{} nor \hydrogen{} network for a cost penalty.  Many of the technologies have not yet been deployed at scale before and will need support for financing, regulation and gaining the acceptance of local populations.
